Deployable Probabilistic Programming

### 2019-04-21 paper/paper.tex
in progress

\linespread{1.1}

@@

\subsection{Model composition}

@@

\subsection{Eight schools}
\label{sec:eight-schools}

@@

\subsection{Linear regression}
\label{sec:linear-regression}

@@

\subsection{Gaussian mixture}
\label{sec:gaussian-mixture}

@@

func (m *Model) Observe(x []float64) float64 {
  ll := 0.0
  mu := make([]float64, m.NComp)
  sigma := make([]float64, m.NComp)

  // Fetch component parameters
  for j := 0; j != m.NComp; j++ {
    mu[j] = x[2*j]
    sigma[j] = math.Exp(x[2*j+1])
  }

  // Compute log likelihood of mixture
  // given the data
  for i := 0; i != len(m.Data); i++ {
    var l float64
    for j := 0; j != m.NComp; j++ {
      lj := Normal.Logp(mu[j], sigma[j],
                        m.Data[i])
      if j == 0 {
        l = lj
      } else {
        l = mathx.LogSumExp(l, lj)
      }
    }
    ll += l
  }
  return ll
}
\end{lstlisting}
\caption{Gaussian mixture: Infergo can handle models with
  dynamic control structure.}
\label{fig:gaussian-mixture}
\end{figure}

>>>>

%\linespread{1.1}

@@

\subsection{Model Composition}

@@

\subsection{Eight Schools}
\label{sec:eight-schools}

@@

\subsection{Linear Regression}
\label{sec:linear-regression}

@@

\subsection{Gaussian Mixture Model}
\label{sec:gaussian-mixture}

@@

func (m *Model) Observe(x []float64) float64 {
  ll := 0.0

  // Impose a prior on component parameters
  ll += Normal.Logps(0, 1, x...)

  // Fetch the parameters
  mu := make([]float64, m.NComp)
  sigma := make([]float64, m.NComp)
  for j := 0; j != m.NComp; j++ {
    mu[j] = x[2*j]
    sigma[j] = math.Exp(x[2*j+1])
  }

  // Compute log likelihood of the mixture
  // given the data
  for i := 0; i != len(m.Data); i++ {
    var l float64
    for j := 0; j != m.NComp; j++ {
      lj := Normal.Logp(mu[j], sigma[j],
                        m.Data[i])
      if j == 0 {
        l = lj
      } else {
        l = logSumExp(l, lj)
      }
    }
    ll += l
  }
  return ll
}

// logSumExp computes log(exp(x)+exp(y)) robustly.
func LogSumExp(x, y float64) float64 {
  z := x
  if y > z {
    z = y
  }
  return z + math.Log(math.Exp(x-z)+math.Exp(y-z))
}

// logSumExp is an elemental, we must register
// the gradient.
func init() {
  ad.RegisterElemental(logSumExp,
    func(_ float64, params ...float64)
      []float64 {
      z := math.Exp(params[1] - params[0])
      t := 1 / (1 + z)
      return []float64{t, t * z}
    })
}
\end{lstlisting}
\caption{Gaussian mixture model: an Infergo model with
  dynamic control structure and a user-defined elemental.}
\label{fig:gaussian-mixture}
\end{figure}
